---
title: "Task 2: Willamette Falls Fish Passage"
author: "German Silva"
date: "3/3/2022"
output: 
  html_document:
    code_folding: hide
    theme: flatly
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# attach packages
library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)
```

# Overview {.tabset .tabset-fade .tabset-pills}

**Summary**:

This report explores adult fish passage data from the Willamette Falls fish ladder on the Willamette River (Oregon, USA) from 2001-01-01 to 2010-12-31. Analysis of the data includes 1) visualization of the original times series, 2) seasonplots for the data, and 3) annual counts by species.    

**Data Source**: 

Columbia River Research, "Data Access in Real Time (DART) Adult Passage Graphics and Text", http://www.cbr.washington.edu/dart/query/adult_graph_text

## Original Time Series

```{r}
adult_pass <- read_csv(here('data', 'willamette_fish_passage.csv')) %>% 
  janitor::clean_names() %>%  
  select(date, coho, steelhead, jack_coho)

adult_pass[is.na(adult_pass)] = 0
```


## Seasonplots

## Annual counts by species


