---
title: 'Task 3: Text Wrangling and Analysis'
author: "Germán Silva"
date: "3/8/2022"
output: 
  html_document:
    theme: flatly
    code_folding: hide
    
---

# Overview:

This report looks at text from the first chapter of the _Wetlands_ textbook by . The text is analyzed to visualize the most common words used throughout the text and text sentiment analysis (is wetland science a positive or negative topic?). 

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach libraries
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
```

# Most Common Words {.tabset .tabset-fade .tabset-pills}

```{r}
# read in the data

wetlands <- pdf_text(here::here("data", "wetlands_chapter_1.pdf"))

# make it a data frame

wetlands_lines <- data.frame(wetlands) %>% # pdf as data frame
  mutate(page = 1:n()) %>% # added page of occurence
  mutate(text_full = str_split(wetlands, pattern = "\\n")) %>% # split lines by line breaks
  unnest(text_full) %>% # unnest lines
  mutate(text_full = str_trim(text_full)) # trim the edges

# get words from lines

wetlands_words <- wetlands_lines %>% # call in the line data
  unnest_tokens(word, text_full) %>% # unnest the text into individual words
  select(-wetlands) 

# clean the word data frame

my_stop <- c("de", "al")

wetlands_words_clean <- wetlands_words %>% # call words
  anti_join(stop_words, by = "word") %>%  # remove stop words (e.g. and, the, a, etc.)
  filter(!str_detect(word, "[0-9]"),
         !str_detect(word, my_stop)) # get rid of numbers 

# get word counts

wetland_clean_counts <- wetlands_words_clean %>% 
  count(word)

# Top 10 words

top_10_words <- wetland_clean_counts %>% 
  arrange(-n) %>% 
  slice(1:10)
  

# Top 100 words

top_100_words <- wetland_clean_counts %>% 
  arrange(-n) %>% 
  slice(1:100)
```

## Top 10 Words Visualization

```{r, fig.align='center', fig.cap= "**Fig. 1** Bar graph showing the top ten words from the text. This is just one way to visualize the such data."}
ggplot(top_10_words, aes(x= reorder(word, -n), y= n))+
  geom_col(fill = "#6d748c")+
  ggtitle("Top 10 Words in CH. 1 of Wetlands")+
  labs(x = "Word",
       y = "Occurence")+
  theme(plot.title = element_text(color = "#5b4f41", hjust = 0.5),
            plot.background = element_rect("white"),
            panel.background = element_rect("#faf7f2"),
            panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
            axis.text = element_text(color = "#5b4f41"),
            axis.title = element_text(color = "#5b4f41"),
            strip.background = element_rect("white"),
            axis.line = element_line(color = "#5b4f41"),
            legend.position = "none")
```

To no surprise, wetlands is the most common word used throughout the text and some other expected high frequency words are in the list such as science, ecology, and science.

**Note**: The way that this analysis is done did not take into account plural words, so wetlands and wetland are counted as two words in the top 10 words.

## Wordcloud

```{r, fig.align='center', fig.cap="**Fig. 2** Another way to visualize text daya is a wordcloud. The above wordcloud shows the top 100 words from the text. Larger words indicate more occurences vs lighter smaller words. "}
ggplot(top_100_words, aes(label = word))+
  geom_text_wordcloud(aes(color = n, size = n), shape = "circle")+
  scale_size_area(max_size = 20)+
  scale_color_gradientn(colors = calecopal::cal_palette(name = "wetland", n= 5, type = "discrete"))+
  theme_minimal()

```

